---
title: "fourDNData"
author: "Jacques Serizay"
date: "`r Sys.Date()`"
output: 
    BiocStyle::html_document
vignette: >
    %\VignetteIndexEntry{fourDNData}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r, eval = TRUE, echo=FALSE, results="hide", message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
suppressPackageStartupMessages({
    library(fourDNData)
})
```

## Access to 4DN data

fourDNData (read *4DN-Data*) is a data package giving programmatic access 
to Hi-C contact matrices uniformly processed by the 
[4DN consortium](https://www.4dnucleome.org/). The matrices are available 
in the multi-resolution `.mcool` format, using the 
`fourDNData(ID = ...)` function. 

```{r}
library(fourDNData)
```

The `fourDNData()` function provides a gateway to ExperimentHub-hosted 4DN 
`.mcool` files.

```{r}
data(fourDNDataFiles)
head(fourDNDataFiles)
cool_file <- fourDNData(id = '4DNFIJTOIGOI')
cool_file
```

## HiCExperiment and fourDNData

`HiCExperiment` package can be used to import `.mcool` files provided by 
`fourDNData`. Refer to `HiCExperiment` package documentation for further 
information. 

```{r}
library(HiCExperiment)
ID <- '4DNFIJTOIGOI'
coolf <- CoolFile(
    path = fourDNData(id = ID), 
    metadata = as.list(fourDNDataFiles[fourDNDataFiles$ID == ID,])
)
x <- import(coolf, resolution = 250000, focus = 'chr5:10000000-50000000')
x
interactions(x)
as(x, 'ContactMatrix')
```

## Session info

```{r}
sessionInfo()
```